---
title: 'ALSM: Chapter 1 Question 47'
subtitle: 'Linear Regression with One Predictor Variable'
author: "Darshan Patel"
date: "12/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(latex2exp)
library(gridExtra)
library(wesanderson)
library(plotly)
```


### Problem 47:
Refer to *Typographical errors* Problem 1.42. Assume that first-order regression model (1.1) is appropriate, with normally distributed independent error terms whose variance is $\sigma^2 = 16$. 

(a) State the likelihood function for the six observations, for $\sigma^2 = 16$.

Answer: Using the model, $Y_i = \beta_0 + \beta_1X_i + \varepsilon_i$, the likelihood function for the six observations is
$$ L(\beta_0, \beta_1) = \prod_{i=1}^6 \frac{1}{\sqrt{2\pi \sigma^2}} \exp\left(-\frac{1}{2\sigma^2}(Y_i - \beta_0 - \beta_1X_i)^2\right) = \prod_{i=1}^6 \frac{1}{\sqrt{32\pi}} \exp\left( -\frac{1}{32}(Y_i - \beta_0 - \beta_1X_i)^2\right) $$ 

(b) Obtain the maximum likelihood estimates of $\beta_0$ and $\beta_1$, using (1.27). 

Answer: Using the maximum likelihood estimation method, the estimator for $\beta_0$ is $\hat{\beta}_0 = b_0$ and for $\beta_1$ is $\hat{\beta}_1 = b_1$. Hence, 
```{r}
typo_errors = read.csv('CH01PR42.txt', sep = '', header = FALSE, 
                  col.names = c('y', 'x'), 
                  colClasses = c('numeric', 'numeric'))

lm.fit_manual = function(X, Y){
  b1 = sum((X - mean(X))*(Y - mean(Y))) / (sum((X - mean(X))^2))
  b0 = mean(Y) - b1*mean(X)
  return(c(b0, b1))
}

typo_mle = lm.fit_manual(typo_errors$x, typo_errors$y)
paste('b0 =', round(typo_mle[1], 3))
paste('b1 =', round(typo_mle[2], 3))
```

(c) Obtain a three-dimensional plot of the likelihood function for values of $\beta_0$ between $\beta_0 = -10$ and $\beta_0 = 10$ and for values of $\beta_1$ between $\beta_1 = 17$ and $\beta_1 = 19$. Does the likelihood appear to be maximized by the maximum likelihood estimates found in part (b)? 

Answer: Create a function for the likelihood function that will take in values of $\beta_0$ and $\beta_1$.
```{r}
typo_mle_fun = function(beta_zero, beta_one){
  return(prod((32*pi)^(-1/2)*exp(-(1/32)*(typo_errors$y - 
                                            beta_zero - 
                                            beta_one*typo_errors$x)^2)))
  }
```

Create a data frame storing the $\beta_0$, $\beta_1$ and likelihood values.
```{r}
n = 350
typo_likelihood_matrix = matrix(rep(NA, n^2), nrow = n, ncol = n)
b0_val = seq(-10, 10, length = n)
b1_val = seq(17, 19, length = n)

for(i in 1:n){
  row = c()
  for(j in b1_val){
    row = c(row, typo_mle_fun(b0_val[i], j))
  }
  typo_likelihood_matrix[i,] = row
}
typo_likelihood_df = data.frame(b0 = b0_val, 
                                typo_likelihood_matrix)
colnames(typo_likelihood_df) = c('b0', b1_val)
typo_likelihood_df = typo_likelihood_df %>% 
  gather(key = b1, value = likelihood, colnames(typo_likelihood_df)[1:n+1])
```

Plot the 3D scatterplot.
```{r}
plot_ly(data = typo_likelihood_df, x = ~b0, y = ~b1, z = ~likelihood, 
        marker = list(color = ~likelihood, colorscale = "RdBu",
                      showscale = TRUE)) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = "beta0"),
                      yaxis = list(title = "beta1"),
                      zaxis = list(title = "likelihood")))
```

By looking at the 3D plot and using the zoom feature, it is apparent that the likelihood is maximized by the maximum likelihood estimates found in part (b).

Using `R`'s `which.max` to find the maximum likelihood value, the following $\beta_0$ and $\beta_1$ estimates create the maximum likelihood value.
```{r}
typo_likelihood_df[which.max(typo_likelihood_df$likelihood),]
```

Same result as before with more precision! 